Nested sampling with two objective functions

### 2015-10-30 Paper/ms.tex
CDF
Include truth figure
Specify the test problem
More words

@@

The following argument is due to John Skilling. Consider Bayesian model selection
between two models $M_1$ and $M_2$. The Bayes Factor is the ratio of the
two marginal likelihoods:
\begin{eqnarray}
B = \frac{P(D | M_1)}{P(D | M_2)} = \frac{Z_1}{Z_2}
\end{eqnarray}
In large problems, the log of the Bayes Factor is usually more convenient:
\begin{eqnarray}
\ln B = \ln Z_1 - \ln Z_2.
\end{eqnarray}
Naively, we want $\ln B$ to an accuracy of about 0.1, so that we can reliably
quantify the strength of the evidence when the two models are similarly
supported by the data. But is this a realistic situation?
Prior to the data, we don't know the
value of $\ln B$. Rather, $\ln B$ has a prior distribution:
\begin{eqnarray}
P(\ln B \in R) &=& \int P(D)P(\ln B \in R | D) \, dD\\
&=& \int P(D) \mathds{1}\left(\ln B(D) \in R\right) \, dD\\
&=& \int \sum_{i\in\{1,2\}}\left[P(M_i)P(D|M_i)\right] \mathds{1}\left(\ln B(D) \in R\right) \, dD
\end{eqnarray}

>>>>

\section{Test problem}

We tested the algorithm on a simple 100-dimensional example, where the
prior is a uniform distribution between 0 and 1 for each coordinate:
\begin{eqnarray}
\pi(\xx) &=& \prod_{i=1}^{100}
\left\{
\begin{array}{lr}
1, & x_i \in [0, 1]\\
0, & \textnormal{otherwise.}
\end{array}
\right.
\end{eqnarray}
The two objective functions are:
\begin{eqnarray}
L_1(\xx) &=& -\sum_{i=1}^{100} \left(x_i - 0.5\right)^2\\
L_2(\xx) &=& -\sum_{i=1}^{100} \log\left[1 +
\left(\frac{x_i - 0.5}{0.01}\right)^2
\right].
\end{eqnarray}
The canonical distributions at low temperature for $L_1$ and high temperature
for $L_2$ are gaussians centered at 0.5 with standard deviations
XX. At high temperatures for $L_2$ and low temperatures for $L_1$ the
canonical distributions are Cauchy centered at 0.5 with width
YY.

\begin{figure}
\centering
\includegraphics[scale=0.75]{figures/truth.pdf}
\caption{\label{fig:truth}}
\end{figure}

@@

The following argument is due to John Skilling. Consider Bayesian model selection
between two models $M_1$ and $M_2$. The Bayes Factor is the ratio of the
two marginal likelihoods:
\begin{eqnarray}
B = \frac{P(D | M_1)}{P(D | M_2)} = \frac{Z_1}{Z_2}
\end{eqnarray}
In large problems, the log of the Bayes Factor is usually more convenient:
\begin{eqnarray}
\ln B = \ln Z_1 - \ln Z_2.
\end{eqnarray}
Naively, we want $\ln B$ to an accuracy of about 0.1, so that we can reliably
quantify the strength of the evidence when the two models are similarly
supported by the data. But is this a realistic situation?
Prior to the data, we don't know the
value of $\ln B$. Rather, $\ln B$ has a prior distribution. We can find this
prior in terms of its CDF:
\begin{eqnarray}
P(\ln B \leq R) &=& \int P(D)P(\ln B \leq R | D) \, dD\\
&=& \int P(D) \mathds{1}\left(\ln B(D) \leq R\right) \, dD\\
&=& \int \sum_{i\in\{1,2\}}\left[P(M_i)P(D|M_i)\right] \mathds{1}\left(\ln B(D) \leq R\right) \, dD
\end{eqnarray}
